---
title: "Create standard variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{calculations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r include=FALSE}
# function to use to display formulas in this vignette.
vignette_equation <- function(.var_name) {
  htmltools::pre(
    gsub("\n  ", "\n", 
         gsub("^  ", "", get_from_calculations_table(.var_name, "equation")))
  )
}

# common text in examples
example1_text <- "All expected variables are provided:"
example2_text <- "All or some of expected variables are missing. In this case, the function will try to detect those missing variables from the parent environment:"
```


The `calculations` family is a collection of functions intended to facilitate the calculation of variables that mostly use standard equations (e.g: CrCl, eGFR, BMI etc). 
This vignette demonstrates: 

* How these functions work when required variables are missing.
* Description of calculations functions.
* Examples use-cases.

```{r, message=FALSE}
library(dmcognigen)
library(dplyr)
```

In their simplest form, these functions apply and report formulas.

```{r}
calculate_crcl(
  age = 50,
  scr = 1,
  sexf = 0,
  wtkg = 70
)
```


# Detect missing variables from environment

When expected variables are missing, an attempt is made to detect them in the environment.

* The calculation function checks the calling environment for missing variables with a case-insensitive match.
* An error is raised if either no match is found or more than one match is found.
* If a single match is found, it is used and the user is notified.


# Source datasets

For examples and testing, there are calculation-ready **sample** datasets incorporated into the package.

## dmcognigen_cov

The `dmcognigen_cov` dataset is a `data.frame` with demographic/covariate information. 
It contains the variables required for most calculation functions. 

```{r}
glimpse(dmcognigen_cov)
```

## dmcognigen_conc

The `dmcognigen_conc` dataset is a `data.frame` with concentration variables.

```{r}
glimpse(dmcognigen_conc)
```


# Functions

Each function performs calculations according to some formula, so the same function can be used to calculate both baseline and time-varying versions of a variable. 
For instance, `calculate_bmi` can calculate both `BMI` (baseline) and `BMIT` (time-varying). 
If you provide time varying `WTKG` and `HTCM`, then the result will be time-varying. 
Each function prints the formula that was used in a readable format.

## calculate_crcl

`r get_from_calculations_table("crcl", "purpose")`

```{r echo=FALSE}
vignette_equation("crcl")
```

`r describe_param("return_num_vect")`. 
Following are the inputs:

* `age`: `r describe_param("age")`
* `wtkg`: `r describe_param("wtkg")`
* `scr`: `r describe_param("scr")`
* `sexf`: `r describe_param("sexf")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    CRCL = calculate_crcl(
      age = AGE,
      scr = SCR,
      sexf = SEXF,
      wtkg = WTKG
    )
  ) %>% 
  select(USUBJID, AGE, SCR, SEXF, WTKG, CRCL)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    CRCL = calculate_crcl()
  ) %>% 
  select(USUBJID, AGE, SCR, SEXF, WTKG, CRCL)

dmcognigen_cov %>% 
  mutate(
    CRCL = calculate_crcl(age = AGE)
  ) %>% 
  select(USUBJID, AGE, SCR, SEXF, WTKG, CRCL)
```


## calculate_crcl_peck

`r get_from_calculations_table("crcl_peck", "purpose")`

```{r echo=FALSE}
vignette_equation("crcl_peck")
```

`r describe_param("return_num_vect")`. Following are the inputs:

* `age`: `r describe_param("age")`
* `crcl`: `r describe_param("crcl")`
* `ibw`: `r describe_param("ibw")`
* `scr`: `r describe_param("scr")`
* `sexf`: `r describe_param("sexf")`
* `wtkg`: `r describe_param("wtkg")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    CRCLP = calculate_crcl_peck(
      age = AGE,
      crcl = CRCL,
      ibw = IBW,
      scr = SCR,
      sexf = SEXF,
      wtkg = WTKG
    )
  ) %>% 
  select(USUBJID, AGE, CRCL, IBW, SCR, SEXF, WTKG, CRCLP)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    CRCLP = calculate_crcl_peck()
  ) %>% 
  select(USUBJID, AGE, CRCL, IBW, SCR, SEXF, WTKG, CRCLP)
```

## calculate_bmi

`r get_from_calculations_table("bmi", "purpose")`

```{r echo=FALSE}
vignette_equation("bmi")
```

`r describe_param("return_num_vect")`. Following are the inputs:

* `htcm`: `r describe_param("htcm")`
* `wtkg`: `r describe_param("wtkg")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    BMI = calculate_bmi(
      htcm = HTCM,
      wtkg = WTKG 
    )
  ) %>% 
  select(USUBJID, HTCM, WTKG, BMI)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    BMI = calculate_bmi()
  ) %>% 
  select(USUBJID, HTCM, WTKG, BMI)
```

## calculate_bsa

`r get_from_calculations_table("bsa", "purpose")`

```{r echo=FALSE}
vignette_equation("bsa")
```

`r describe_param("return_num_vect")`. Following are the inputs:

* `htcm`: `r describe_param("htcm")`
* `wtkg`: `r describe_param("wtkg")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    BSA = calculate_bsa(
      htcm = HTCM, 
      wtkg = WTKG
    )
  ) %>% 
  select(USUBJID, HTCM, WTKG, BSA)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    BSA = calculate_bsa()
  ) %>% 
  select(USUBJID, HTCM, WTKG, BSA)
```

## calculate_lbm

`r get_from_calculations_table("lbm", "purpose")`

```{r echo=FALSE}
vignette_equation("lbm")
```

`r describe_param("return_num_vect")`. Following are the inputs:

* `htcm`: `r describe_param("htcm")`
* `sexf`: `r describe_param("sexf")`
* `wtkg`: `r describe_param("wtkg")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    LBM = calculate_lbm(
      htcm = HTCM,
      sexf = SEXF,
      wtkg = WTKG
    )
  ) %>% 
  select(USUBJID, HTCM, SEXF, WTKG, LBM)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    LBM = calculate_lbm()
  ) %>% 
  select(USUBJID, HTCM, SEXF, WTKG, LBM)
```

## calculate_ibw

`r get_from_calculations_table("ibw", "purpose")`

```{r echo=FALSE}
vignette_equation("ibw")
```

`r describe_param("return_num_vect")`. Following are the inputs:

* `htcm`: `r describe_param("htcm")`
* `sexf`: `r describe_param("sexf")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    IBW = calculate_ibw(
      htcm = HTCM,
      sexf = SEXF 
    )
  ) %>% 
  select(USUBJID, HTCM, SEXF, IBW)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    IBW = calculate_ibw()
  ) %>% 
  select(USUBJID, HTCM, SEXF, IBW)
```

## calculate_ibw_child

`r get_from_calculations_table("ibw_child", "purpose")`

```{r echo=FALSE}
vignette_equation("ibw_child")
```

`r describe_param("return_num_vect")`. Following is the input:

* `age`: `r describe_param("age")`
* `htcm`: `r describe_param("htcm")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    IBWCHILD = calculate_ibw_child(age = AGE, htcm = HTCM)
  ) %>% 
  select(USUBJID, AGE, HTCM, IBWCHILD)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    IBWCHILD = calculate_ibw_child()
  ) %>% 
  select(USUBJID, AGE, HTCM, IBWCHILD)
```

## calculate_egfr

`r get_from_calculations_table("egfr", "purpose")`

```{r echo=FALSE}
vignette_equation("egfr")
```

`r describe_param("return_num_vect")`. Following are the inputs:

* `age`: `r describe_param("age")`
* `racen`: `r describe_param("racen")`
* `scr`: `r describe_param("scr")`
* `sexf`: `r describe_param("sexf")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    EGFR = calculate_egfr(
      age = AGE,
      racen = RACEN,
      scr = SCR,
      sexf = SEXF
    )
  ) %>% 
  select(USUBJID, AGE, RACEN, SCR, SEXF, EGFR)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    EGFR = calculate_egfr()
  ) %>% 
  select(USUBJID, AGE, RACEN, SCR, SEXF, EGFR)
```

## calculate_egfr_child

`r get_from_calculations_table("egfr_child", "purpose")`

```{r echo=FALSE}
vignette_equation("egfr_child")
```

`r describe_param("return_num_vect")`. Following are the inputs:

* `age`: `r describe_param("age")`
* `htcm`: `r describe_param("htcm")`
* `scr`: `r describe_param("scr")`
* `sexf`: `r describe_param("sexf")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    EGFR_CHILD = calculate_egfr_child(
      age = AGE, 
      htcm = HTCM,
      scr = SCR,
      sexf = SEXF
    )
  ) %>% 
  select(USUBJID, AGE, HTCM, SCR, SEXF, EGFR_CHILD)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    EGFR_CHILD = calculate_egfr_child()
  ) %>% 
  select(USUBJID, AGE, HTCM, SCR, SEXF, EGFR_CHILD)
```

## calculate_rfcat

`r get_from_calculations_table("rfcat", "purpose")`

```{r echo=FALSE}
vignette_equation("rfcat")
```

`r describe_param("return_num_vect")`. Following is the input:

* `egfr`: `r describe_param("egfr")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    EGFR = calculate_egfr(
      age = AGE,
      racen = RACEN,
      scr = SCR,
      sexf = SEXF
    ),
    RFCAT = calculate_rfcat(egfr = EGFR)
  ) %>% 
  select(USUBJID, EGFR, RFCAT)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    RFCAT = calculate_rfcat()
  ) %>% 
  select(USUBJID, EGFR, RFCAT)
```

## calculate_tbilcat

`r get_from_calculations_table("tbilcat", "purpose")`

```{r echo=FALSE}
vignette_equation("tbilcat")
```

`r describe_param("return_num_vect")`. Following are the inputs:

* `tbil`: `r describe_param("tbil")`
* `tbiluln`: `r describe_param("uln")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    TBILCAT = calculate_tbilcat(
      tbil = TBIL, 
      tbiluln = TBILULN
    )
  ) %>% 
  select(USUBJID, TBIL, TBILULN, TBILCAT)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    TBILCAT = calculate_tbilcat()
  ) %>% 
  select(USUBJID, TBIL, TBILULN, TBILCAT)
```

## calculate_astcat

`r get_from_calculations_table("astcat", "purpose")`

```{r echo=FALSE}
vignette_equation("astcat")
```

`r describe_param("return_num_vect")`. Following are the inputs:

* `ast`: `r describe_param("ast")`
* `astuln`: `r describe_param("uln")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    ASTCAT = calculate_astcat(
      ast = AST, 
      astuln = ASTULN
    )
  ) %>% 
  select(USUBJID, AST, ASTULN, ASTCAT)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    ASTCAT = calculate_astcat()
  ) %>% 
  select(USUBJID, AST, ASTULN, ASTCAT)
```

## calculate_nciliv

`r get_from_calculations_table("nciliv", "purpose")`

```{r echo=FALSE}
vignette_equation("nciliv")
```

`r describe_param("return_num_vect")`. Following are the inputs:

* `astcat`: `r describe_param("astcat")`
* `tbilcat`: `r describe_param("tbilcat")`

### Examples

`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    NCILIV = calculate_nciliv(astcat = ASTCAT, tbilcat = TBILCAT)
  ) %>% 
  select(USUBJID, ASTCAT, TBILCAT, NCILIV)
```

`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    NCILIV = calculate_nciliv()
  ) %>% 
  select(USUBJID, ASTCAT, TBILCAT, NCILIV)
```
