---
title: "Create standard variables with dmcognigen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{calculations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `calculations` family is a collection of functions intended to facilitate the calculation of variables that mostly use standard equations (e.g: CrCl, eGFR, BMI etc). This vignette shows and explains: 

* How these functions work when expected variables are not passed.
* The sample data integrated in the package, and how you can use it to practice.
* Description of calculations functions.
* Examples from different use cases.

# Loading `calculations`
By installing `dmcognigen` package and loading it into your environment, you will have access to the `calculations` functions.

Assuming you already have `dmcognigen` package installed on your machine, then you load it into your environment:
```{r, message=FALSE}
library(dmcognigen)
library(dplyr)
```


# Detect missing variables from environment
There's a common behavior between all calculations functions when all or some of expected variables are missing. They will try to detect and use those missing variables from environment, details below: 

* It identifies the missing variables/inputs, if found, then
* It looks in the environment to see if it contains those missing variables (it's a case-insensitive match)
* It'll stop and pass error if either no match found or more than one matches found
* If only one match found, then it'll be used in calculation and user will be notified


# Source datasets
In order to provide some examples and facilitating the testing for users, we've integrated some calculation ready **sample** datasets into the package. Once you've loaded the package, you can simply access them by calling their names.

```{r}
dmcognigen_cov %>%
  mutate(
    CRCL = calculate_crcl(
      age = AGE,
      wtkg = WTKG,
      scr = SCR,
      sexf = SEXF
    )
  ) %>%
  select(USUBJID, AGE, WTKG, SCR, SEXF, CRCL)
```

## dmcognigen_cov
The `dmcognigen_cov` dataset is a `data.frame` type. It contains the necessary variables to test and practice the calculation functions. Some variables like SCR are already created in this dataset, so you don't need to create it.

```{r}
glimpse(dmcognigen_cov)
```

## dmcognigen_conc
The `dmcognigen_conc` dataset is a `data.frame` type. It contains the concentration variables like DV to facilitate testing and practicing.

```{r}
glimpse(dmcognigen_conc)
```


<!-- This (Stationary vs Time varying variables) is explained in # Function section.  -->

# Functions
Each function calculates a specific variable, but the same function can be used to calculate both baseline and time varying of a variable because it depends on the inputs. For instance, `calculate_bmi` can calculate both `BMI` and `BMIT`. If you provide time varying `WTKG` and `HTCM`, then the output will be a time varying `BMI`. Additionally, each function prints the equation that was used to calculate in a readable format.

## calculate_crcl
The `calculate_crcl` function is to `r get_from_calculations_table("crcl","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("crcl","equation"))`</pre> `r describe_param("return_num_vect")`. Following are the inputs:

* `age`: `r describe_param("age")`
* `wtkg`: `r describe_param("wtkg")`
* `scr`: `r describe_param("scr")`
* `sexf`: `r describe_param("sexf")`

### Examples

<!-- Common text between sections -->
```{r, include = FALSE}
example1_text <- "All expected variables are provided:"

example2_text <- "All or some of expected variables are missing. In this case, the function will try to detect those missing variables from the parent environment:"
```

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    CRCL = calculate_crcl(
      age = AGE,
      wtkg = WTKG,
      scr = SCR,
      sexf = SEXF
    )
  ) %>% 
  select(USUBJID, AGE, WTKG, SCR, SEXF, CRCL)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    CRCL = calculate_crcl()
  ) %>% 
  select(USUBJID, AGE, WTKG, SCR, SEXF, CRCL)

dmcognigen_cov %>% 
  mutate(
    CRCL = calculate_crcl(age = AGE)
  ) %>% 
  select(USUBJID, AGE, WTKG, SCR, SEXF, CRCL)
```


## calculate_crcl_peck
The `calculate_crcl_peck` function is to `r get_from_calculations_table("crcl_peck","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("crcl_peck","equation"))`</pre> `r describe_param("return_num_vect")`. Following are the inputs:

* `age`: `r describe_param("age")`
* `wtkg`: `r describe_param("wtkg")`
* `crcl`: `r describe_param("crcl")`
* `ibw`: `r describe_param("ibw")`
* `scr`: `r describe_param("scr")`
* `sexf`: `r describe_param("sexf")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    CRCLP = calculate_crcl_peck(
      age = AGE,
      wtkg = WTKG,
      crcl = CRCL,
      ibw = IBW,
      scr = SCR,
      sexf = SEXF
    )
  ) %>% 
  select(USUBJID, AGE, WTKG, CRCL, IBW, SCR, SEXF, CRCLP)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    CRCLP = calculate_crcl_peck()
  ) %>% 
  select(USUBJID, AGE, WTKG, CRCL, IBW, SCR, SEXF, CRCLP)
```

## calculate_bmi
The `calculate_bmi` function is to `r get_from_calculations_table("bmi","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("bmi","equation"))`</pre> `r describe_param("return_num_vect")`. Following are the inputs:

* `wtkg`: `r describe_param("wtkg")`
* `htcm`: `r describe_param("htcm")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    BMI = calculate_bmi(
      wtkg = WTKG, 
      htcm = HTCM
    )
  ) %>% 
  select(USUBJID, WTKG, HTCM, BMI)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    BMI = calculate_bmi()
  ) %>% 
  select(USUBJID, WTKG, HTCM, BMI)
```

## calculate_bsa
The `calculate_bsa` function is to `r get_from_calculations_table("bsa","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("bsa","equation"))`</pre> `r describe_param("return_num_vect")`. Following are the inputs:

* `htcm`: `r describe_param("htcm")`
* `wtkg`: `r describe_param("wtkg")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    BSA = calculate_bsa(
      htcm = HTCM, 
      wtkg = WTKG
    )
  ) %>% 
  select(USUBJID, HTCM, WTKG, BSA)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    BSA = calculate_bsa()
  ) %>% 
  select(USUBJID, HTCM, WTKG, BSA)
```

## calculate_lbm
The `calculate_lbm` function is to `r get_from_calculations_table("lbm","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("lbm","equation"))`</pre> `r describe_param("return_num_vect")`. Following are the inputs:

* `sexf`: `r describe_param("sexf")`
* `wtkg`: `r describe_param("wtkg")`
* `htcm`: `r describe_param("htcm")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    LBM = calculate_lbm(
      sexf = SEXF,
      wtkg = WTKG,
      htcm = HTCM
    )
  ) %>% 
  select(USUBJID, SEXF, WTKG, HTCM, LBM)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    LBM = calculate_lbm()
  ) %>% 
  select(USUBJID, SEXF, WTKG, HTCM, LBM)
```

## calculate_ibw
The `calculate_ibw` function is to `r get_from_calculations_table("ibw","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("ibw","equation"))`</pre> `r describe_param("return_num_vect")`. Following are the inputs:

* `sexf`: `r describe_param("sexf")`
* `htcm`: `r describe_param("htcm")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    IBW = calculate_ibw(
      sexf = SEXF, 
      htcm = HTCM
    )
  ) %>% 
  select(USUBJID, SEXF, HTCM, IBW)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    IBW = calculate_ibw()
  ) %>% 
  select(USUBJID, SEXF, HTCM, IBW)
```

## calculate_ibw_child
The `calculate_ibw_child` function is to `r get_from_calculations_table("ibw_child","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("ibw_child","equation"))`</pre> `r describe_param("return_num_vect")`. Following is the input:

* `htcm`: `r describe_param("htcm")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    IBWCHILD = calculate_ibw_child(htcm = HTCM)
  ) %>% 
  select(USUBJID, HTCM, IBWCHILD)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    IBWCHILD = calculate_ibw_child()
  ) %>% 
  select(USUBJID, HTCM, IBWCHILD)
```

## calculate_egfr
The `calculate_egfr` function is to `r get_from_calculations_table("egfr","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("egfr","equation"))`</pre> `r describe_param("return_num_vect")`. Following are the inputs:

* `scr`: `r describe_param("scr")`
* `age`: `r describe_param("age")`
* `sexf`: `r describe_param("sexf")`
* `racen`: `r describe_param("racen")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    EGFR = calculate_egfr(
      scr = SCR,
      age = AGE,
      sexf = SEXF,
      racen = RACEN
    )
  ) %>% 
  select(USUBJID, SCR, AGE, SEXF, RACEN, EGFR)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    EGFR = calculate_egfr()
  ) %>% 
  select(USUBJID, SCR, AGE, SEXF, RACEN, EGFR)
```

## calculate_egfr_child
The `calculate_egfr_child` function is to `r get_from_calculations_table("egfr_child","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("egfr_child","equation"))`</pre> `r describe_param("return_num_vect")`. Following are the inputs:

* `htcm`: `r describe_param("htcm")`
* `scr`: `r describe_param("scr")`
* `age`: `r describe_param("age")`
* `sexf`: `r describe_param("sexf")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    EGFR_CHILD = calculate_egfr_child(
      htcm = HTCM,
      scr = SCR,
      age = AGE, 
      sexf = SEXF
    )
  ) %>% 
  select(USUBJID, HTCM, SCR, AGE, SEXF, EGFR_CHILD)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    EGFR_CHILD = calculate_egfr_child()
  ) %>% 
  select(USUBJID, HTCM, SCR, AGE, SEXF, EGFR_CHILD)
```

## calculate_rfcat
The `calculate_rfcat` function is to `r get_from_calculations_table("rfcat","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("rfcat","equation"))`</pre> `r describe_param("return_num_vect")`. Following is the input:

* `egfr`: `r describe_param("egfr")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    EGFR = calculate_egfr(
      scr = SCR,
      age = AGE,
      sexf = SEXF,
      racen = RACEN
    ),
    RFCAT = calculate_rfcat(egfr = EGFR)
  ) %>% 
  select(USUBJID, EGFR, RFCAT)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    RFCAT = calculate_rfcat()
  ) %>% 
  select(USUBJID, EGFR, RFCAT)
```

## calculate_tbilcat
The `calculate_tbilcat` function is to `r get_from_calculations_table("tbilcat","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("tbilcat","equation"))`</pre> `r describe_param("return_num_vect")`. Following are the inputs:

* `tbil`: `r describe_param("tbil")`
* `tbiluln`: `r describe_param("uln")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    TBILCAT = calculate_tbilcat(
      tbil = TBIL, 
      tbiluln = TBILULN
    )
  ) %>% 
  select(USUBJID, TBIL, TBILULN, TBILCAT)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    TBILCAT = calculate_tbilcat()
  ) %>% 
  select(USUBJID, TBIL, TBILULN, TBILCAT)
```

## calculate_astcat
The `calculate_astcat` function is to `r get_from_calculations_table("astcat","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("astcat","equation"))`</pre> `r describe_param("return_num_vect")`. Following are the inputs:

* `ast`: `r describe_param("ast")`
* `astuln`: `r describe_param("uln")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    ASTCAT = calculate_astcat(
      ast = AST, 
      astuln = ASTULN
    )
  ) %>% 
  select(USUBJID, AST, ASTULN, ASTCAT)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    ASTCAT = calculate_astcat()
  ) %>% 
  select(USUBJID, AST, ASTULN, ASTCAT)
```

## calculate_nciliv
The `calculate_nciliv` function is to `r get_from_calculations_table("nciliv","purpose")`. Following is the formula to calculate it: <pre>`r gsub("  ", "", get_from_calculations_table("nciliv","equation"))`</pre> `r describe_param("return_num_vect")`. Following are the inputs:

* `tbilcat`: `r describe_param("tbilcat")`
* `astcat`: `r describe_param("astcat")`

### Examples

#### Example 1
`r example1_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    NCILIV = calculate_nciliv(tbilcat = TBILCAT, astcat = ASTCAT)
  ) %>% 
  select(USUBJID, TBILCAT, ASTCAT, NCILIV)
```

#### Example 2
`r example2_text`
```{r}
dmcognigen_cov %>% 
  mutate(
    NCILIV = calculate_nciliv()
  ) %>% 
  select(USUBJID, TBILCAT, ASTCAT, NCILIV)
```