---
title: "Decode Tables"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Decode Tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(dmcognigen)
library(dplyr)

data("dmcognigen_pk_requirements")
requirements <- dmcognigen_pk_requirements %>% 
  select(variable_name, variable_label, format_decode)

data("dmcognigen_cov")
cov <- dmcognigen_cov

data("dmcognigen_pk")
pk <- dmcognigen_pk
```


# Introduction

Decode tables (`decode_tbls`) are defined based on variable names and their levels and labels. 
The level is generally a shorthand representation of the label and is ideally numeric. 
The label is intended to be a more detailed description, or the text to display on outputs like graphs and tables. 
These `decode_tbls` can be used to provide specifications to functions like `set_decode_factors()`, `join_decode_labels()`, and `join_decode_levels()`.

For example, below is a decode table for a variable `RFCAT` (`r attr(cov$RFCAT, "label")`). 
The values on the left-hand-side of the equal sign are the levels and the values on the right-hand-side are the labels.
```{r echo=FALSE}
requirements %>% 
  filter(variable_name == "RFCAT") %>% 
  with(extract_decode_tbl(
    variable_name = variable_name,
    decode = format_decode
  ))
```


# Constructing `decode_tbls`

A decode table (`decode_tbl`) or a collection of decode tables (`decode_tbls`) can be constructed in 3 main ways:

1. Based on a character string that describes the levels and labels.
1. Based on variables in data.
1. By manually constructing data frame(s).

While `decode_tbls` can be defined many ways, the primary intention is that `read_requirements()` is used to define this and other attributes.

## Extracting from character strings

This method is expected to be used when variables and their decodes are presented together in a table and is used in `read_requirements()` to define the `"decode_tbls"` attribute.

Like for the below variable descriptions:
```{r echo=FALSE}
requirements %>% 
  filter(format_decode != "")
```

For this method, the decodes are expected to be of type `character`.
Each entry on a new line defines one level-to-label relationship separated by an equal sign.
When an entry contains more than one equal sign, the first equal sign is considered the separator.
```{r echo=FALSE}
# print the variable names and the content of their decode
requirements %>% 
  filter(format_decode != "") %>% 
  rowwise() %>% 
  summarise(cat(paste0(
    variable_name, ":\n", format_decode, "\n\n"
  ))) %>% 
  invisible()
```

To extract the decode tables from these types of strings:
```{r}
extract_decode_tbls(
  variable_name = requirements$variable_name,
  decode = requirements$format_decode
)
```


## Extracting from a dataset

This method is expected to be used when a data set already includes level and label variables.

Like `RACEN` and `RACEC` below:
```{r}
cov %>%
  cnt(RACEN, RACEC)

cov %>%
  extract_decode_tbls_from_data(
    lvl_to_lbl = list(RACEN = "RACEC")
  )
```

The `lvl_to_lbl` argument maps the names of level variables to the names of label variables.
```{r}
cov %>%
  extract_decode_tbls_from_data(
    lvl_to_lbl = list(
      # map individual variables
      SEXF = "SEXFC",
      RACEN = "RACEC",
      # map other lvl to lbl by removing CD at the end of variable names
      ~ stringr::str_remove(.x, "CD$")
    )
  )
```

Sometimes, more than one representation of a variable is in a data set. 
Like how this `cov` data set includes the pair of variables `RACEN` & `RACEC`, along with the original source variable `RACE`.
```{r}
cov %>%
  extract_decode_tbls_from_data(
    lvl_to_lbl = list(
      RACEN = "RACE"
    )
  )
```

Ideally, the level variable is numeric. 
But other data types are accepted. 
One way to review merged content would be to map a label variable to the original source variable it was derived from.
```{r}
cov %>%
  extract_decode_tbls_from_data(
    lvl_to_lbl = list(
      RACE = "RACEC"
    )
  )
```

## Constructing manually from a data frame

This can be done many different ways. 
In general, use `as_decode_tbls()` with a named list of data frames that contain variables `var`, `lvl`, and `lbl`.
One example is below, where decodes are defined for multiple variables, then the named list is generated with `split()`.

```{r}
tibble::tribble(
  ~var, ~lvl, ~lbl,
  "RACEN", 1, "White/Caucasian",
  "RACEN", 2, "Black/African American",
  "RACEN", 3, "Asian",
  "RACEN", 4, "American Indian or Alaska Native",
  
  "SEXF",  0, "Male",
  "SEXF",  1, "Female"
  ) %>% 
  split(~ var) %>% 
  as_decode_tbls()
```


# Incorporating `decode_tbls` as variables in data sets

To demonstrate the automation features of these utilities, consider a data set that contains only numeric variables. 
```{r}
pk_numeric <- pk %>% 
  select(where(is.numeric))
```

## `set_decode_factors()`

Modify or create new factor variables based on `decode_tbls` or `requirements` objects. 
The order of the levels is defined based on the sort order of the level in the decode definition. 
This is useful for other functions that consider the order of factor levels.
```{r}
pk_numeric %>% 
  set_decode_factors(requirements) %>% 
  cnt(RACEN, SEXF, n_distinct_vars = ID)
```

Since the resulting variables are factors, they are easy to summarize with `across()`. 
This example provides a summary of categorical covariates by `ID`.
```{r}
pk_numeric %>% 
  select(all_of(stationary_variables(., ID))) %>% 
  set_decode_factors(requirements) %>% 
  cnt(across(where(is.factor)), n_distinct_vars = ID)
```

The `new_names` argument works similarly to the `lvl_to_lbl` argument in joining functions, but mapping an existing variable to itself is allowed by `set_decode_factors()`.
```{r}
pk_numeric %>% 
  set_decode_factors(
    decode_tbls = requirements, 
    new_names = list(
      "{var}FCT",
      RACEN = "RACEN", 
      SEXF = "SEXFC" 
    )
  ) %>% 
  cnt(RACEN, SEXFC, RFCATFCT, n_distinct_vars = ID)
```


## `join_decode_labels()`

Create new label variables for variables with matching names in data and `decode_tbls`.
```{r}
pk_numeric_with_label_vars <- pk_numeric %>% 
  join_decode_labels(requirements)
```

The default label name will be the level name appended with a C. 
Or, use `lvl_to_lbl` to provide a named list that can include other glue specifications or functions to map the level names to the label names. 
One unnamed element can be included in `lvl_to_lbl` to provide default behavior.
```{r}
pk_numeric %>% 
  select(ID, RACEN, SEXF) %>% 
  join_decode_labels(requirements, lvl_to_lbl = list(RACEN = "RACEC", "{var}C")) %>% 
  cnt(RACEN, RACEC, SEXF, SEXFC, n_distinct_vars = ID)
```

Variables will not be joined if they already exist.
```{r}
pk %>% 
  select(ID, RACEN, RACEC, SEXF) %>% 
  join_decode_labels(requirements, lvl_to_lbl = list(RACEN = "RACEC", "{var}C")) %>% 
  cnt(RACEN, RACEC, SEXF, SEXFC, n_distinct_vars = ID)
```

## `join_decode_levels()`

For scenarios where a label variable is already defined in a data set, the corresponding levels can be joined with `join_decode_levels()`.

The `lvl_to_lbl` structure is identical whether joining labels or levels.
```{r}
pk %>% 
  select(ID, RACEC, SEXFC) %>% 
  join_decode_levels(requirements, lvl_to_lbl = list(RACEN = "RACEC", "{var}C")) %>% 
  cnt(RACEN, RACEC, SEXF, SEXFC, n_distinct_vars = ID)
```

