---
title: "Decodes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Decodes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dmcognigen)
library(dplyr)
```

```{r echo=FALSE}
# copied content from vignettes/requirements.Rmd
reqs_to_write <- tibble::tribble(
  ~variable_name, ~variable_label,             ~pk_ard, ~pk_mif, ~format_decode, 
  "STUDYID",      "Study Identifier",          "x",     NA,      NA,
  "USUBJID",      "Unique Subject Identifier", "x",     NA,      NA,
  "ID",           "Subject ID",                "x",     "x",     NA,
  "AGECAT",       "Baseline Age Category",     "x",     "x",     "1=18-50\n2=51-69\n3=70+",
  "RACEN",        "Race",                      "x",     "x",     "1=White/Caucasian\n2=Black/African American\n3=Asian-Japanese\n4=Asian-Korean\n5=Asian-other\n6=American Indian or Alaska Native\n7=Native Hawaiian or Other Pacific Islander\n8=Other",
  "SEXF",         "Sex",                       "x",     "x",     "0=Male\n1=Female",
  "FED",          "Fed",                       "x",     "x",     "0=Fasted\n1=Fed"
)

example_decodes <- reqs_to_write %>% 
  select(variable_name, variable_label, format_decode)
```

# Introduction

Decodes are defined by a variable along with its levels and labels. The level is generally a shorthand (or numeric) representation of the label. For example, below is a decode table for a variable AGECAT. The values on the left-hand-side are the levels and the values on the right-hand-side are the labels.
```{r echo=FALSE}
example_decodes %>% 
  filter(variable_name == "AGECAT") %>% 
  with(extract_decode_tbl(
    variable_name = variable_name,
    decode = format_decode
  ))
```

# Constructing decode tables

A decode table (`decode_tbl`) or a collection of decode tables (`decode_tbls`) can be constructed in 2 main ways:

1. Based on a character string that describes the levels and labels.
1. Based on variables in data.

## Extracting from character strings

This method is expected to be used when variables and their decodes are presented together in a table, like the below table:
```{r echo=FALSE}
example_decodes
```

The decodes themselves include one or more lines, where each line defines a level-to-label relationship separated by an equal sign.
```{r echo=FALSE}
# print the variable names and the content of their decode
example_decodes %>% 
  filter(!is.na(format_decode)) %>% 
  rowwise() %>% 
  summarise(cat(paste0(
    variable_name, ":\n", format_decode, "\n\n"
  ))) %>% 
  invisible()
```

