---
title: "Decodes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Decodes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dmcognigen)
library(dplyr)
```

```{r echo=FALSE}
example_decodes <- dmcognigen:::example_requirements %>% 
  select(variable_name, variable_label, format_decode)
```

# Introduction

Decodes are defined by a variable along with its levels and labels. The level is generally a shorthand (or numeric) representation of the label. For example, below is a decode table for a variable AGECAT. The values on the left-hand-side are the levels and the values on the right-hand-side are the labels.
```{r echo=FALSE}
example_decodes %>% 
  filter(variable_name == "AGECAT") %>% 
  with(extract_decode_tbl(
    variable_name = variable_name,
    decode = format_decode
  ))
```

# Constructing decode tables

A decode table (`decode_tbl`) or a collection of decode tables (`decode_tbls`) can be constructed in 2 main ways:

1. Based on a character string that describes the levels and labels.
1. Based on variables in data.

## Extracting from character strings

This method is expected to be used when variables and their decodes are presented together in a table, like the below table:
```{r echo=FALSE}
example_decodes
```

The decodes themselves include one or more lines, where each line defines a level-to-label relationship separated by an equal sign.
```{r echo=FALSE}
# print the variable names and the content of their decode
example_decodes %>% 
  filter(!is.na(format_decode)) %>% 
  rowwise() %>% 
  summarise(cat(paste0(
    variable_name, ":\n", format_decode, "\n\n"
  ))) %>% 
  invisible()
```

We can extract the decode tables as below:
```{r}
extract_decode_tbls(
  variable_name = example_decodes$variable_name,
  decode = example_decodes$format_decode
)
```


## Extracting from a dataset

This method is expected to be used when a dataset includes variables and their decodes.

Like RACEN and RACE below:
```{r}
dmcognigen_cov %>%
  cnt(RACEN, RACE)

dmcognigen_cov %>%
  extract_decode_tbls_from_data(
    lvl_to_lbl = c("RACEN" = "RACE")
  )
```

The `lvl_to_lbl` argument maps the names of "level" variables to the names of "label" variables.
```{r}
dmcognigen_cov %>%
  extract_decode_tbls_from_data(
    lvl_to_lbl = c(
      # map individual variables
      "SEXF" = "SEXFC",
      "RACEN" = "RACE",
      # by default map lvl to lbl by removing CD at the end of variable names
      ~ stringr::str_remove(.x, "CD$")
    )
  )
```

