---
title: "Data Requirements"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Requirements}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dmcognigen)
library(dplyr)

data("dmcognigen_pk_requirements")

data("dmcognigen_cov")
data("dmcognigen_pk")
```

# Introduction

Data requirements are specifications of a dataset.
They describe required variables with varying details like variable label, type, description, source, formula, rounding specifications, and relationships to other variables. 

Generally, the requirements are maintained in a file. 
These functions support gathering and applying this information from Excel files, Word (docx) tables, or a `data.frame`. 

For the purposes of this vignette, consider files in a directory that include tables with these variables:
```{r echo=FALSE}
tibble::tribble(
  ~name,             ~description,
  "variable_name",   "Variable Name",
  "variable_label",  "Variable Label",
  "pk_ard",          "Whether the variable is in the Analysis Ready Dataset",
  "pk_mif",          "Whether the variable is in the Model Input File",
  "format_decode",   "Pairs of (generally) numeric values and descriptions"
) %>% 
  print.data.frame(row.names = FALSE)
```

# Identify data requirements files

With a versioning strategy based on the latest modification date of the file, there might be multiple versions of the data requirements file that exist in the data assembly directory. 

Consider the data assembly directory below:
```{r}
asmbdat_directory <- file.path(tempdir(), "asmbdat")
asmbdat_directory
```

```{r include=FALSE}
# write the file to tmp a few times, to demonstrate the capture of the date in the filename
dir.create(asmbdat_directory, showWarnings = FALSE)

# pretend that a long time ago, we didn't name the specs tab
openxlsx::write.xlsx(
  dmcognigen_pk_requirements,
  file = file.path(asmbdat_directory, "data-requirements-2019-12-01.xlsx")
)

openxlsx::write.xlsx(
  list(specs = dmcognigen_pk_requirements),
  file = file.path(asmbdat_directory, "data-requirements-2020-01-01.xlsx")
)

# include additional sheet (empty)
openxlsx::write.xlsx(
  list(specs = dmcognigen_pk_requirements, unit_conversions = data.frame()),
  file = file.path(asmbdat_directory, "data-requirements-2020-01-10.xlsx")
)

# include another additional sheet (empty)
# this is intended to be the latest version.
openxlsx::write.xlsx(
  list(specs = dmcognigen_pk_requirements, unit_conversions = data.frame(), discussions = data.frame()),
  file = file.path(asmbdat_directory, paste0("data-requirements-", lubridate::today(), ".xlsx"))
)

# copy of the requirements for QC purposes
openxlsx::write.xlsx(
  list(specs = dmcognigen_pk_requirements, unit_conversions = data.frame(), discussions = data.frame(), qc_findings = data.frame()),
  file = file.path(asmbdat_directory, paste0("qc-data-requirements-", lubridate::today(), ".xlsx"))
)

# empty docx file to demonstrate including docx in searches
file.create(file.path(asmbdat_directory, "legacy-data-requirements-2000-01-01.docx"))
```

```{r echo=FALSE}
fs::dir_tree(asmbdat_directory)
```

Use the `available_requirements_table()` function to collect information about multiple data requirements files. 
Supported files that match the `pattern` and have the required `sheet` names or positions will be kept. 
By default, the required sheet name is `"specs"`.

The resulting table can be used to identify the desired version of the data requirements file to pass to the `read_requirements()` function. 

```{r}
available_requirements_table(
  path = asmbdat_directory,
  pattern = "req"
) %>% 
  select(-c(created, modified)) %>% 
  print.data.frame()
```

Use `sheet = NULL` for no required sheet names. 
This is required to include docx files in these searches.
```{r}
available_requirements_table(
  path = asmbdat_directory,
  pattern = "req",
  sheet = NULL
) %>%
  select(-c(created, modified)) %>% 
  print.data.frame()
```

Use `drop_qc = FALSE` to include versions of the data requirements used for QC. 
These files are ignored otherwise.
```{r}
available_requirements_table(
  path = asmbdat_directory,
  pattern = "req",
  sheet = NULL,
  drop_qc = FALSE
) %>% 
  select(-c(created, modified)) %>% 
  print.data.frame()
```

# Read data requirements

If no `path` is provided, `read_requirements()` will read the latest version of the matching data requirements files in the working directory. 
The latest version is selected as the file with the most recent date in the filename. 
If no dates are detected in the filenames, the matching file with the most recent modification time is selected. 
```{r}
requirements <- read_requirements(asmbdat_directory)
glimpse(requirements)
```

Use the `subset` argument to provide subset/filter criteria:
```{r}
requirements_pk_mif <- read_requirements(
  asmbdat_directory,
  subset = pk_mif == "x"
)
glimpse(requirements_pk_mif)
```

Use `as_requirements()` to apply requirements attributes to an existing data frame:
```{r}
requirements_pk_mif_with_as <- as_requirements(
  as.data.frame(requirements_pk_mif)
)
identical(requirements_pk_mif_with_as, requirements_pk_mif)
```


## Attributes

When possible, additional attributes are assigned to the result of `read_requirements()` and `as_requirements()`. 
Resulting `requirements` objects can be used directly in other functions in this package to apply these attributes. 

### decode_tbls

This attribute will be set if the `variable_name_col` and `decode_col` arguments are defined and those variables are in requirements. 

See more details on using `decode_tbls` in the [Decode Tables vignette](decodes.html).

```{r}
attr(requirements, "decode_tbls")
```


#### Join decodes

To join new variables to a data set based on `decode_tbls` or `requirements` objects, use `join_decode_labels()` or `join_decode_levels()`.
```{r}
dmcognigen_pk %>% 
  select(USUBJID, RACEN, SEXF) %>% 
  join_decode_labels(dmcognigen_pk_requirements, lvl_to_lbl = list(RACEN = "RACEC", "{var}C")) %>% 
  cnt(RACEN, RACEC, SEXF, SEXFC, n_distinct_vars = USUBJID)
```


#### Create or modify variables as factors

Use `set_decode_factors()` to either modify variables in-place or create new variables. 
```{r}
dmcognigen_cov %>% 
  set_decode_factors(requirements, new_names = list(RACEN = "RACEN", "{var}FCT")) %>% 
  cnt(RACEN, across(ends_with("FCT")), n_cumulative = FALSE)
```


### labels_named_list

This attribute will be set if the `variable_name_col` and `variable_label_col` arguments are defined and those variables are in the `requirements`. 
The `requirements` can be used directly in `set_labels()` or the attribute can be extracted and possibly modified.
```{r}
attr(requirements, "labels_named_list") %>% 
  head()
```

```{r include=FALSE}
empty_labels <- character(ncol(dmcognigen_cov)) %>% 
  purrr::set_names(names(dmcognigen_cov))

dmcognigen_cov_labels_removed <- dmcognigen_cov %>% 
  set_labels(labels = empty_labels) %>% 
  select(STUDYID, USUBJID, RACEN)
```

Apply the labels to a data set:
```{r}
dmcognigen_cov_labels_removed %>% 
  str()

# apply the labels
dmcognigen_cov_labels_removed %>% 
  set_labels(labels = requirements) %>% 
  str()
```
