---
title: "Leverage data requirements"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Leverage data requirements}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dmcognigen)
library(dplyr)
```

# Introduction

Data requirements are specifications of a desired dataset, describing expected variables, their labels, and their content. Generally, we document the requirements for a dataset within an Excel file where one of the sheets provides information about desired variables.  Other sheets contain a variety of supporting information.

For the purposes of this vignette, consider an Excel file that includes a sheet with the below table:
```{r echo=FALSE}
reqs_to_write <- dmcognigen:::example_requirements

reqs_to_write
```

Where the variables names indicate:
```{r echo=FALSE}
tibble::tribble(
  ~name,             ~description,
  "variable_name",   "Variable Name",
  "variable_label",  "Variable Label",
  "pk_ard",          "Whether to keep the variable in the Analysis Ready Dataset",
  "pk_mif",          "Whether to keep the variable in the Model Input File",
  "format_decode",   "Name/Value pairs of decodes, also referred to as formats"
) %>% 
  print.data.frame()
```


# Identify data requirements files

With a versioning strategy based on the latest modification date of the file, there might be multiple versions of the data requirements file that exist in the data assembly directory. 

Consider the data assembly directory below:
```{r}
asmbdat_directory <- file.path(tempdir(), "asmbdat")
asmbdat_directory
```

```{r include=FALSE}
# write the file to tmp a few times, to demonstrate the capture of the date in the filename
dir.create(asmbdat_directory, showWarnings = FALSE)

# pretend that a long time ago, we didn't name the specs tab
openxlsx::write.xlsx(
  reqs_to_write,
  file = file.path(asmbdat_directory, "data-requirements-2019-12-01.xlsx")
)

openxlsx::write.xlsx(
  list(specs = reqs_to_write),
  file = file.path(asmbdat_directory, "data-requirements-2020-01-01.xlsx")
)

# include additional sheet (empty)
openxlsx::write.xlsx(
  list(specs = reqs_to_write, unit_conversions = data.frame()),
  file = file.path(asmbdat_directory, "data-requirements-2020-01-10.xlsx")
)

# include another additional sheet (empty)
openxlsx::write.xlsx(
  list(specs = reqs_to_write, unit_conversions = data.frame(), discussions = data.frame()),
  file = file.path(asmbdat_directory, "data-requirements-2020-01-20.xlsx")
)

# copy of the requirements for QC purposes
openxlsx::write.xlsx(
  list(specs = reqs_to_write, unit_conversions = data.frame(), discussions = data.frame(), qc_findings = data.frame()),
  file = file.path(asmbdat_directory, "qc-data-requirements-2020-01-20.xlsx")
)
```

```{r echo=FALSE}
# list the files, preferably with fs::dir_tree
if(requireNamespace("fs", quietly = TRUE)) {
  fs::dir_tree(asmbdat_directory)
} else {
  list.files(asmbdat_directory)
}
```

Use the `available_requirements_table()` function to collect information about multiple data requirements files. Excel files that match the `pattern` and have the required `sheet` names or positions will be kept. By default, the required sheet name is `"specs"`.
```{r}
available_requirements_table(
  path = asmbdat_directory,
  pattern = "req"
) %>% 
  print.data.frame()
```

Use `sheet = NULL` for no required sheet names.
```{r}
available_requirements_table(
  path = asmbdat_directory,
  pattern = "req",
  sheet = NULL
) %>% 
  print.data.frame()
```

Use `drop_qc = FALSE` to include versions of the data requirements used for QC. These files are ignored otherwise.
```{r}
available_requirements_table(
  path = asmbdat_directory,
  pattern = "req",
  sheet = NULL,
  drop_qc = FALSE
) %>% 
  print.data.frame()
```

The resulting table can be used to identify the desired version of the data requirements file to pass to the `read_requirements()` function. 

# Read data requirements

By default, `read_requirements()` will read the latest version of the matching data requirements files in the working directory. The latest version is selected as the file with the most recent date in the filename. If no dates are detected in the filenames, the matching file with the most recent modification time is selected.
```{r}
requirements <- read_requirements(asmbdat_directory)
glimpse(requirements)
```

Use the `subset` argument to provide subset/filter criteria:
```{r}
requirements_pk_mif <- read_requirements(
  asmbdat_directory,
  subset = pk_mif == "x"
)
glimpse(requirements_pk_mif)
```

## Attributes

When possible, additional attributes are assigned to the result of `read_requirements()`. These attributes have use-cases for various stages of data assembly.

### decode_tbls

This attribute will be set if the `variable_name_col` and `decode_col` arguments are defined and variables with those names exist in the imported requirements. 

```{r}
attr(requirements, "decode_tbls")
```

Or the `decode_tbls` can can be extracted:
```{r}
extracted_requirements_decode_tbls <- extract_decode_tbls(
  variable_name = requirements$variable_name, 
  decode = requirements$format_decode
)
identical(attr(requirements, "decode_tbls"), extracted_requirements_decode_tbls)
```

This attribute can be used to create new variables based on shared names between the data and `decode_tbls`.
```{r}
# shared variables
variables_with_decodes <- intersect(names(extracted_requirements_decode_tbls), names(dmcognigen_cov))
variables_with_decodes
```

The default new name will be the original name appended with a C. Or, use `lvl_to_lbl` to provide a named list that can include other glue specifications or functions to map the level names to the label names. One unnamed element can be included in `lvl_to_lbl` to provide default behavior.
```{r}
dmcognigen_cov %>% 
  select(USUBJID, RACEN, SEXF) %>% 
  join_decode_labels(requirements, lvl_to_lbl = list(RACEN = "RACEC", "{var}C")) %>% 
  cnt(RACEN, RACEC, SEXF, SEXFC)
```

### labels_named_list

This attribute will be set if the `variable_name_col` and `variable_label_col` arguments are defined and variables with those names exist in the imported requirements. 

```{r}
attr(requirements, "labels_named_list")
```

```{r include=FALSE}
empty_labels <- character(ncol(dmcognigen_cov)) %>% 
  purrr::set_names(names(dmcognigen_cov))

dmcognigen_cov_labels_removed <- dmcognigen_cov %>% 
  set_labels(labels = empty_labels) %>% 
  select(STUDYID, USUBJID, RACEN)
```

This attribute can be used in the `set_labels()` function, or the imported requirements can be used:
```{r}
dmcognigen_cov_labels_removed %>% 
  str()

# apply the labels
dmcognigen_cov_labels_removed %>% 
  set_labels(labels = requirements) %>% 
  str()
```

