---
title: "Search Environment Data"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Search Environment Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(dmcognigen)
library(tidyverse)
```

`search_environment_data()` searches for patterns in variable names, variable labels, and variable content.
`cnt_search_result()` summarizes the matching results.
These functions are particularly useful when the source of some information is expected to exist, but is unknown.
The default `envir` is the global environment, but other environments and named lists that include data frames are also accepted.

To demonstrate, load all of the `pharmaversesdtm` datasets into an environment.
```{r}
pharmaversesdtm_data_env <- new.env()

data(
  list = data(package = "pharmaversesdtm")$results[ , "Item"], 
  package = "pharmaversesdtm", 
  envir = pharmaversesdtm_data_env
)
```

Then search for a pattern:
```{r}
search_environment_data(
  regex("XANOMELINE", ignore_case = TRUE),
  envir = pharmaversesdtm_data_env
)
```

To immediately review the results that matched the pattern, pass to `cnt_search_result()`.
```{r}
search_environment_data(
  regex("XANOMELINE", ignore_case = TRUE),
  envir = pharmaversesdtm_data_env
) %>% 
  cnt_search_result(
    n_distinct_vars = "USUBJID"
  )
```

One way to further collect and review the results is to `bind_rows()`.
```{r}
search_environment_data(
  regex("STUDYID", ignore_case = TRUE),
  envir = pharmaversesdtm_data_env
) %>% 
  cnt_search_result(
    n_distinct_vars = "USUBJID"
  ) %>% 
  bind_rows(.id = "dataset")
```

Optional arguments can be used to include/exclude datasets/variables.
```{r}
search_environment_data(
  regex("cancer", ignore_case = TRUE),
  envir = pharmaversesdtm_data_env
) %>% 
  cnt_search_result(
    n_distinct_vars = "USUBJID",
    ignore_df_names = c("ae", "ae_ophtha"),
    extra_vars = c("STUDYID")
  )
```
